---
title: Doing a Sentiment Analysis on Tweets (Part 2)
author: Jacob Rozran
date: '2014-12-29'
slug: doing-a-sentiment-analysis-on-tweets-part-2
categories:
  - Data Engineering
  - Data Analysis
tags:
  - data engineering
  - data analysis
  - R
  - twitteR
  - sentiment analysis
---

### INTRO

This is post is a continuation of 
[my last post](https://www.jakelearnsdatascience.com/posts/doing-a-sentiment-analysis-on-tweets-part-1/). 
There I pulled tweets from Twitter related to “Comcast email,” got rid of the junk, 
and removed the unnecessary/unwanted data.

Now that I have the tweets, I will further clean the text and subject it to two 
different analyses: emotion and polarity.

### WHY DOES THIS MATTER

Before I get started, I thought it might be a good idea to talk about WHY I am 
doing this (besides the fact that I learned a new skill and want to show it off 
and get feedback). This yet incomplete project was devised for two reasons:

1. Understand the overall customer sentiment about the product I support
2. Create an early warning system to help identify when things are going wrong on 
the platform

Keeping the customer voice at the forefront of everything we do is tantamount to 
providing the best experience for the users of our platform. Identifying trends 
in sentiment and emotion can help inform the team in many ways, including seeing 
the reaction to new features/releases (i.e. – seeing a rise in comments about a 
specific addition from a release) and identifying needed changes to current 
functionality (i.e. – users who continually comment about a specific behavior of 
the application) and improvements to user experience (i.e. – trends in comments 
about being unable to find a certain feature on the site). Secondarily, this 
analysis can act as an early warning system when there are issues with the 
platform (i.e. – a sudden spike in comments about the usability of a mobile 
device).

Now that I’ve explained why I am doing this (which I probably should have done 
in this sort of detail the first post), let’s get into how it is actually done…

### STEP ONE: STRIPPING THE TEXT FOR ANALYSIS

There are a number of things included in tweets that dont matter for the analysis. 
Things like twitter handles, URLs, punctuation… they are not necessary to do the 
analysis (in fact, they may well confound it). This bit of code handles that 
cleanup.

For those following the scripts on GitHub, this is part of my 
[tweet_clean.R script](https://github.com/jrozra200/scraping_twitter/blob/master/tweet_clean.R). 
Also, to give credit where it is due: I’ve borrowed and tweaked the code from 
[Andy Bromberg’s blog](http://andybromberg.com/sentiment-analysis/) to do this task.

```{r tweet_clean, eval = FALSE}
library(stringr) ##Does some of the text editing

##Cleaning up the data some more (just the text now) First grabbing only the text
text <- paredTweetList$Tweet

# remove retweet entities
text <- gsub("(RT|via)((?:\\b\\W*@\\w+)+)", "", text)
# remove at people
text <- gsub("@\\w+", "", text)
# remove punctuation
text <- gsub("[[:punct:]]", "", text)
# remove numbers
text <- gsub("[[:digit:]]", "", text)
# remove html links
text <- gsub("http\\w+", "", text)

# define "tolower error handling" function
try.error <- function(x) {
# create missing value
y <- NA
# tryCatch error
try_error <- tryCatch(tolower(x), error=function(e) e)
# if not an error
if (!inherits(try_error, "error"))
y <- tolower(x)
# result
return(y)
}
# lower case using try.error with sapply
text <- sapply(text, try.error)

# remove NAs in text
text <- text[!is.na(text)]
# remove column names
names(text) <- NULL
```

### STEP TWO: CLASSIFYING THE EMOTION FOR EACH TWEET

So now the text is just that: only text. The punctuation, links, handles, etc. 
have been removed. Now it is time to estimate the emotion of each tweet.

Through some research, I found that there are many posts/sites on Sentiment 
Analysis/Emotion Classification that use the 
[“Sentiment” package in R](https://cran.r-project.org/web/packages/sentiment/index.html). 
I thought: “Oh great – a package tailor made to solve the problem for which I 
want an answer.” The problem is that this package has been deprecated and removed 
from the CRAN library.

To get around this, I downloaded the archived package and pulled the code for 
doing the emotion classification. With some minor tweaks, I was able to get it 
going. This can be seen in its entirety in the 
[classify_emotion.R script](https://github.com/jrozra200/scraping_twitter/blob/master/classify_emotion.R).
 You can also see the “made for the internet” version here:
 
``` {r classify_emotion, eval = FALSE}
library(RTextTools)
library(tm)
algorithm <- "bayes"
prior <- 1.0
verbose <- FALSE
matrix <- create_matrix(text)
lexicon <- read.csv("./data/emotions.csv.gz",header=FALSE)

counts <- list(anger=length(which(lexicon[,2]=="anger")), 
               disgust=length(which(lexicon[,2]=="disgust")), 
               fear=length(which(lexicon[,2]=="fear")), 
               joy=length(which(lexicon[,2]=="joy")), 
               sadness=length(which(lexicon[,2]=="sadness")), 
               surprise=length(which(lexicon[,2]=="surprise")), 
               total=nrow(lexicon))
documents <- c()

for (i in 1:nrow(matrix)) {
        if (verbose) print(paste("DOCUMENT",i))
        scores <- list(anger=0,disgust=0,fear=0,joy=0,sadness=0,surprise=0)
        doc <- matrix[i,]
        words <- findFreqTerms(doc,lowfreq=1)

        for (word in words) {
                for (key in names(scores)) {
                        emotions <- lexicon[which(lexicon[,2]==key),]
                        index 0) {
                                entry <- emotions[index,]
                                
                                category <- as.character(entry[[2]]])
                                count <- counts[[category]]

                                score <- 1.0
                                if (algorithm=="bayes") score <- abs(log(score*prior/count))
                                if (verbose) {
                                        print(paste("WORD:",word,"CAT:",
                                                    category,"SCORE:",score))
                                        }
                                
                                scores[[category]] <- scores[[category]]+score
                                }
                        }
                }
        
        if (algorithm=="bayes") {
                for (key in names(scores)) {
                        count <- counts[[key]]
                        total <- counts[["total"]]
                        score <- abs(log(count/total))
                        scores[[key]] <- scores[[key]]+score
                        }
                } else {
                        for (key in names(scores)) {
                                scores[[key]] <- scores[[key]]+0.000001
                                }
                        }
        
        best_fit <- names(scores)[which.max(unlist(scores))]
        if (best_fit == "disgust" && as.numeric(unlist(scores[2]))-3.09234 < .01) best_fit <- NA
        documents <- rbind(documents, c(scores$anger, 
                                        scores$disgust, 
                                        scores$fear, 
                                        scores$joy, 
                                        scores$sadness, 
                                        scores$surprise, 
                                        best_fit))
        }

colnames(documents) <- c("ANGER", "DISGUST", "FEAR", "JOY", 
                         "SADNESS", "SURPRISE", "BEST_FIT")
```

Here is a sample output from this code:

| ANGER | DISGUST | FEAR | JOY | SADNESS | SURPRISE | BEST_FIT |
|-------|---------|------|-----|---------|----------|----------|
| “1.46871776464786” | “3.09234031207392” | “2.06783599555953” | “1.02547755260094” | “7.34083555412328” | “7.34083555412327” | “sadness” |
| “7.34083555412328” | “3.09234031207392” | “2.06783599555953” | “1.02547755260094” | “1.7277074477352” | “2.78695866252273” | “anger” |
| “1.46871776464786” | “3.09234031207392” | “2.06783599555953” | “1.02547755260094” | “7.34083555412328” | “7.34083555412328” | “sadness” |

Here you can see that the initial author is using naive Bayes (which honestly I 
don’t yet understand) to analyze the text. I wanted to show a quick snipet of 
how the analysis is being done “under the hood.”

For my purposes though, I only care about the emotion outputted and the tweet 
it is analyzed from.

```{r snippet, eval = FALSE}
emotion <- documents[, "BEST_FIT"]`
```

This variable, emotion, is returned by the classify_emotion.R script.

#### CHALLENGES OBSERVED

In addition to not fully understanding the code, the emotion classification 
seems to only work OK (which is pretty much expected… this is a canned analysis 
that hasn’t been tailored to my analysis at all). I’d like to come back to this 
one day to see if I can do a better job analyzing the emotions of the tweets.

### STEP THREE: CLASSIFYING THE POLARITY OF EACH TWEET

Similarly to what we saw in step 5, I will use the cleaned text to analyze the 
polarity of each tweet.

This code is also from the old R Packaged titled “Sentiment.” As with above, I 
was able to get the code working with only some minor tweaks. This can be seen 
in its entirety in the 
[classify_polarity.R script](https://github.com/jrozra200/scraping_twitter/blob/master/classify_polarity.R). 
Here it is, too:

```{r classify_polarity, eval = FALSE}
algorithm <- "bayes"
pstrong <- 0.5
pweak <- 1.0
prior <- 1.0
verbose <- FALSE
matrix <- create_matrix(text)
lexicon <- read.csv("./data/subjectivity.csv.gz",header=FALSE)
counts <- list(positive=length(which(lexicon[,3]=="positive")),
               negative=length(which(lexicon[,3]=="negative")), 
               total=nrow(lexicon))
documents <- c()

for (i in 1:nrow(matrix)) {
        if (verbose) print(paste("DOCUMENT",i))
        scores <- list(positive=0,negative=0)
        doc <- matrix[i,]
        words <- findFreqTerms(doc, lowfreq=1)
        
        for (word in words) {
                index 0) {
                        entry <- lexicon[index,]
                        
                        polarity <- as.character(entry[[2]])
                        category <- as.character(entry[[3]])
                        count <- counts[[category]]
                        
                        score <- pweak
                        if (polarity == "strongsubj") score <- pstrong
                        if (algorithm=="bayes") score <- abs(log(score*prior/count))
                        
                        if (verbose) {
                                print(paste("WORD:", word, "CAT:", 
                                            category, "POL:", polarity, 
                                            "SCORE:", score))
                                }
                        
                        scores[[category]] <- scores[[category]]+score
                }
        }
        
        if (algorithm=="bayes") {
                for (key in names(scores)) {
                        count <- counts[[key]]
                        total <- counts[["total"]]
                        score <- abs(log(count/total))
                        scores[[key]] <- scores[[key]]+score
                        }
                } else {
                        for (key in names(scores)) {
                                scores[[key]] <- scores[[key]]+0.000001
                        }
                }
        
        best_fit <- names(scores)[which.max(unlist(scores))]
        ratio <- as.integer(abs(scores$positive/scores$negative))
        if (ratio==1) best_fit <- "neutral"
        documents <- rbind(documents,c(scores$positive, 
                                       scores$negative, 
                                       abs(scores$positive/scores$negative), 
                                       best_fit))
        
        if (verbose) {
                print(paste("POS:", scores$positive,"NEG:", 
                            scores$negative, "RATIO:", 
                            abs(scores$positive/scores$negative)))
                cat("\n")
                }
        }

colnames(documents) <- c("POS","NEG","POS/NEG","BEST_FIT")
```

Here is a sample output from this code:

| POS | NEG | POS/NEG | BEST_FIT |
|-----|-----|---------|----------|
| “1.03127774142571” | “0.445453222112551” | “2.31512017476245” | “positive” |
| “1.03127774142571” | “26.1492093145274” | “0.0394381997949273” | “negative” |
| “17.9196623384892” | “17.8123396772424” | “1.00602518608961” | “neutral” |

Again, I just wanted to show a quick snipet of how the analysis is being done 
“under the hood.”

I only care about the polarity outputted and the tweet it is analyzed from.

```{r snippet_two, eval = FALSE}
polarity <- documents[, "BEST_FIT"]
```

This variable, polarity, is returned by the classify_polarity.R script.

#### CHALLENGES OBSERVED

As with above, this is a stock analysis and hasn’t been tweaked for my needs. 
The analysis does OK, but I want to come back to this again one day to see if 
I can do better.

### QUICK CONCLUSION

So… Now I have the emotion and polarity for each tweet. This can be useful to 
see on its own, but I think is more worthwhile in aggregate. In my next post, 
I’ll show that.

Also in the next post- I’ll also show an analysis of the word count with a 
wordcloud… This gets into the secondary point of this analysis. Hypothetically, 
I’d like to see common issues bubbled up through the wordcloud.