---
title: Sierpinski Triangles (and Carpets) in R
author: Jacob Rozran
date: '2015-10-05'
slug: sierpinski-triangles-and-carpets-in-r
categories:
  - Sierpinski Triangles
tags:
  - learning
  - sierpinski triangles
  - stat theory
  - R
  - RStudio
---

Recently in class, I was asked the following question:

> Start with an equilateral triangle and a point chosen at random from the interior of that triangle. Label one vertex 1, 2, a second vertex 3, 4, and the last vertex 5, 6. Roll a die to pick a vertex. Place a dot at the point halfway between the roll-selected vertex and the point you chose. Now consider this new dot as a starting point to do this experiment once again. Roll the die to pick a new vertex. Place a dot at the point halfway between the last point and the most recent roll-selected vertex. Continue this procedure. What does the shape of the collection of dots look like?

I thought, well - it's got to be something cool or else the professor wouldn't ask, but I can't imagine it will be more than a cloud of dots. Truth be told, I went to a conference for work the week of this assignment and never did it - but when I went to the next class, **IT WAS SOMETHING COOL!** It turns out that this creates a [Sierpinski Triangle](https://en.wikipedia.org/wiki/Sierpinski_triangle) - a fractal of increasingly smaller triangles.

I wanted to check this out for myself, so I built an R script that creates the 
triangle. I ran it a few times with differing amounts of points. Here is one 
with 50,000 points. Though this post is written in RStudio, I've hidden the code 
for readability. Actual code for this can be found 
[here](https://github.com/jrozra200/pretty_triangles/blob/master/createTriangle.R).

```{r sierpinski_triangle, warning = FALSE, message = FALSE, out.width = "100%", echo = FALSE}
## CREATE THE TRIANGLE WITH SIDES OF LENGTH a, CREATING n POINTS, AND        ##
## STARTING POINT AT (startx, starty)                                        ##
createTriange <- function(a = 10, n = 50000, startx = a / 2, 
                          starty = sqrt(3) * a / 4){
        
        ## ESTABLISH THE CORNERS OF THE TRIANGLE                             ##
        ## corner1 IS THE FIRST POINT AT THE ORIGIN; corner2 IS THE SECOND   ##
        ## POINT AT A DISTANCE OF a FROM corner1 AND HALFWAY (ON THE X AXIS) ##
        ## BETWEEN corner1 AND corner3; corner3 IS ON THE X AXIS A DISTANCE  ##
        ## OF a FROM corner1 and corner2                                     ##
        corner1 <- data.frame(x = 0, y = 0) 
        corner2 <- data.frame(x = a / 2, y = sqrt(3) * a / 2)
        corner3 <- data.frame(x = a, y = 0) 
        
        ## CREATE THE STARTING POINT                                         ##
        starting_point <- data.frame(x = startx, y = starty)
        
        ## ADD THE ESTABLISHED POINTS TO THE all_points DATA FRAME - THESE   ##
        ## WILL BE PLOTTED WILL ALL OF THE OTHER POINTS THAT ARE CREATED     ##
        all_points <- data.frame(starting_point)
        
        ## LOOP TO CREATE n POINTS                                           ##
        for(i in 1:n){
                ## CREATE A RANDOM NUMBER BETWEEN 1 AND 4 (ONE FOR EACH      ##
                ## CORNER)                                                   ##
                rand_num <- sample(1:3, 1)
                
                ## PICK A CORNER TO GO TOWARDS BASED ON THE RANDOM NUMBER    ##
                tmp_corn <- if(rand_num == 1){
                        corner1
                } else if(rand_num == 2){
                        corner2
                } else {
                        corner3
                }
                
                ## FIND THE DISTANCE THAT IS HALFWAY BETWEEN THE CORNER AND  ##
                ## POINT WE ARE CURRENTLY AT                                 ##
                tmp_dist <- sqrt((all_points$x[i] - tmp_corn$x)^2 + (all_points$y[i] - tmp_corn$y)^2) / 2
                
                ## FIND THE SLOPE FROM THE CURRENT POINT TO THE CORNER POINT ##
                tmp_slope <- (all_points$y[i] - tmp_corn$y) / (all_points$x[i] - tmp_corn$x)
                
                ## CREATE THE DIVIDEND FOR THE EQUATIONS BELOW               ##
                tmp_r <- sqrt(1 + (tmp_slope)^2)
                
                ## FIND THE NEW VALUE OF X (ADDING OR SUBTRACTING DEPENDING  ##
                ## ON WHERE THE POINT IS IN REGARDS TO WHERE IT IS GOING)    ##
                x <- if(all_points$x[i] < tmp_corn$x) {
                        all_points$x[i] + (tmp_dist / tmp_r)
                } else if(all_points$x[i] > tmp_corn$x){
                        all_points$x[i] - (tmp_dist / tmp_r)
                } else {
                        all_points$x[i]
                }
                
                ## FIND THE NEW VALUE OF Y (ADDING OR SUBTRACTING DEPENDING  ##
                ## ON WHERE THE POINT IS IN REGARDS TO WHERE IT IS GOING)    ##
                y <- if ((all_points$y[i] >= tmp_corn$y) & (all_points$x[i] 
                                                            >= tmp_corn$x)){
                        all_points$y[i] - (tmp_dist * tmp_slope / tmp_r)
                } else if ((all_points$y[i] >= tmp_corn$y) & (all_points$x[i] 
                                                              < tmp_corn$x)){
                        all_points$y[i] + (tmp_dist * tmp_slope / tmp_r)
                } else if ((all_points$y[i] < tmp_corn$y) & (all_points$x[i] 
                                                             >= tmp_corn$x)){
                        all_points$y[i] - (tmp_dist * tmp_slope / tmp_r)        
                } else {
                        all_points$y[i] + (tmp_dist * tmp_slope / tmp_r)
                }
                
                ## PUT X AND Y TOGETHER IN THE NEW POINT                     ##
                tmp_point <- c(x, y)
                
                ## ADD THE NEW POINT TO THE FRAME OF ALL POINTS              ##
                all_points <- rbind(all_points, tmp_point)
        }
        
        ## ADD THE CORNERS TO THE LIST OF POINTS TO BE PLOTTED               ##
        all_points <- rbind(all_points, corner1, corner2, corner3)
        
        ## PLOT THE POINTS                                                   ##
        plot(all_points$x, all_points$y, pch = ".", xlab = "", ylab = "")
}

createTriange()
```

I thought - if equilateral triangles create patterns this cool, a square must be 
amazing! Well... it is, however you can't just run this logic - it will return a 
cloud of random dots... 

After talking with my professor, Dr. Levitan - it turns out you can get something 
equally awesome as the Sierpinski triangle with a square; you just need to make 
a few changes (say this with a voice of authority and calm knowingness):

> Instead of 3 points to move to, you need 8 points: the 4 corners of a specified square and the midpoints between each side. Also, instead of taking the midpoint of your move to the specified location, you need to take the tripoint (division by 3 instead of 2).

This is called a [Sierpinski Carpet](https://en.wikipedia.org/wiki/Sierpinski_carpet) 
- a fractal of squares (as opposed to a fractal of equilateral triangles in the 
graph above). You can see in both the triangle and square that the same pattern 
is repeated time and again in smaller and smaller increments.

I updated my R script and voila - MORE BEAUTIFUL MATH! 

``` {r sierpinski_carpet, warning = FALSE, message = FALSE, out.width = "100%", echo = FALSE}
## CREATE THE SQUARE WITH SIDES OF LENGTH a, CREATING n POINTS, AND          ##
## STARTING POINT AT (startx, starty)                                        ##
createSquare <- function(a = 10, n = 50000, startx = a / 2, 
                          starty = a / 2){
        
        ## ESTABLISH THE CORNERS OF THE SQUARE                               ##
        ## corner1 IS THE FIRST POINT AT THE ORIGIN; corner2 IS THE SECOND   ##
        ## POINT AT (0, a), corner3 IS THE THIRD POINT AT (a, a), AND        ##
        ## CORNER 4 IS THE LAST POINT AT (a, 0)                              ##
        corner1 <- data.frame(x = 0, y = 0) 
        corner2 <- data.frame(x = 0, y = a)
        corner3 <- data.frame(x = a, y = a) 
        corner4 <- data.frame(x = a, y = 0)
        midpoint1 <- data.frame(x = 0, y = a / 2)
        midpoint2 <- data.frame(x = a / 2, y = a)
        midpoint3 <- data.frame(x = a, y = a / 2)
        midpoint4 <- data.frame(x = a / 2, y = 0)
        
        ## CREATE THE STARTING POINT                                         ##
        starting_point <- data.frame(x = startx, y = starty)
        
        ## ADD THE ESTABLISHED POINTS TO THE all_points DATA FRAME - THESE   ##
        ## WILL BE PLOTTED WILL ALL OF THE OTHER POINTS THAT ARE CREATED     ##
        all_points <- data.frame(starting_point)
        
        ## LOOP TO CREATE n POINTS                                           ##
        for(i in 1:n){
                ## CREATE A RANDOM NUMBER BETWEEN 1 AND 4 (ONE FOR EACH      ##
                ## CORNER)                                                   ##
                rand_num <- sample(1:8, 1)
                
                ## PICK A CORNER TO GO TOWARDS BASED ON THE RANDOM NUMBER    ##
                tmp_corn <- if (rand_num == 1) { 
                        corner1
                } else if (rand_num == 2) {
                        corner2
                } else if (rand_num == 3) {
                        corner3
                } else if (rand_num == 4) {
                        corner4
                } else if (rand_num == 5) {
                        midpoint1
                } else if (rand_num == 6) {
                        midpoint2
                } else if (rand_num == 7) {
                        midpoint3
                } else{
                        midpoint4
                }
                
                ## FIND THE DISTANCE THAT IS HALFWAY BETWEEN THE CORNER AND  ##
                ## POINT WE ARE CURRENTLY AT                                 ##
                tmp_dist <- sqrt((all_points$x[i] - tmp_corn$x)^2 + (all_points$y[i] - tmp_corn$y)^2) * (2 / 3)
                
                ## FIND THE SLOPE FROM THE CURRENT POINT TO THE CORNER POINT ##
                tmp_slope <- (all_points$y[i] - tmp_corn$y) / (all_points$x[i] - tmp_corn$x)
                
                ## CREATE THE DIVIDEND FOR THE EQUATIONS BELOW               ##
                tmp_r <- sqrt(1 + (tmp_slope)^2)
                
                ## FIND THE NEW VALUE OF X (ADDING OR SUBTRACTING DEPENDING  ##
                ## ON WHERE THE POINT IS IN REGARDS TO WHERE IT IS GOING)    ##
                x <- if(all_points$x[i] < tmp_corn$x) {
                        all_points$x[i] + (tmp_dist / tmp_r)
                } else if(all_points$x[i] > tmp_corn$x){
                        all_points$x[i] - (tmp_dist / tmp_r)
                } else {
                        all_points$x[i]
                }
                
                ## FIND THE NEW VALUE OF Y (ADDING OR SUBTRACTING DEPENDING  ##
                ## ON WHERE THE POINT IS IN REGARDS TO WHERE IT IS GOING)    ##
                y <- if ((all_points$y[i] >= tmp_corn$y) & (all_points$x[i] 
                                                            >= tmp_corn$x)){
                        all_points$y[i] - (tmp_dist * tmp_slope / tmp_r)
                } else if ((all_points$y[i] >= tmp_corn$y) & (all_points$x[i] 
                                                              < tmp_corn$x)){
                        all_points$y[i] + (tmp_dist * tmp_slope / tmp_r)
                } else if ((all_points$y[i] < tmp_corn$y) & (all_points$x[i] 
                                                             >= tmp_corn$x)){
                        all_points$y[i] - (tmp_dist * tmp_slope / tmp_r)        
                } else {
                        all_points$y[i] + (tmp_dist * tmp_slope / tmp_r)
                }
                
                ## PUT X AND Y TOGETHER IN THE NEW POINT                     ##
                tmp_point <- c(x, y)
                
                ## ADD THE NEW POINT TO THE FRAME OF ALL POINTS              ##
                all_points <- rbind(all_points, tmp_point)
        }
        
        ## ADD THE CORNERS TO THE LIST OF POINTS TO BE PLOTTED               ##
        all_points <- rbind(all_points, corner1, corner2, corner3, corner4,
                            midpoint1, midpoint2, midpoint3, midpoint4)
        
        ## PLOT THE POINTS                                                   ##
        plot(all_points$x, all_points$y, pch = ".", xlab = "", ylab = "")
}

createSquare()
```

Check out the script and run the functions yourself! I only spent a little bit of time 
putting it together - I think it would be cool to add some other features, 
especially when it comes to the plotting of the points. Also - I'd like to run 
it for a million or more points... I just lacked the patience to wait out the 
script to run for that long (50,000 points took about 30 minutes to run - my 
script is probably not the most efficient).

Anyways - really cool to see what happens in math sometimes - its hard to imagine 
at first that the triangle would look that way. Another reason math is cool!