---
title: Doing a Sentiment Analysis on Tweets (Part 1)
author: Jacob Rozran
date: '2014-12-24'
slug: doing-a-sentiment-analysis-on-tweets-part-1
categories:
  - Data Engineering
  - Data Analysis
tags:
  - data engineering
  - data analysis
  - R
  - twitteR
  - sentiment analysis
---

### INTRO

So... This post is my first foray into the R 
[twitteR](https://cran.r-project.org/web/packages/twitteR/index.html) package. 
This post assumes that you have that package installed already in R. I show here 
how to get tweets from Twitter in preparation for doing some sentiment analysis. 
My next post will be the actual sentiment analysis.

For this example, I am grabbing tweets related to "Comcast email." My goal of 
this exercise is to see how people are feeling about the product I support.

### STEP 1: GETTING AUTHENTICATED TO TWITTER

First, you'll need to create an application at Twitter. I used 
[this blog post](http://thinktostart.com/twitter-authentification-with-r/) to 
get rolling with that. This post does a good job walking you through the steps 
to do that.

Once you have your app created, this is the code I used to create and save 
my authentication credentials. Once you've done this once, you need only load 
your credentials in the future to authenticate with Twitter.

```{r twitteR_setup, eval = FALSE}
library(twitteR) ## R package that does some of the Twitter API heavy lifting

consumerKey <- "INSERT YOUR KEY HERE"
consumerSecret <- "INSERT YOUR SECRET HERE"
reqURL <- "https://api.twitter.com/oauth/request_token "
accessURL <- "https://api.twitter.com/oauth/access_token "
authURL <- "https://api.twitter.com/oauth/authorize "
twitCred <- OAuthFactory$new(consumerKey = consumerKey,
                             consumerSecret = consumerSecret,
                             requestURL = reqURL,
                             accessURL = accessURL,
                             authURL = authURL)

twitCred$handshake()

save(cred, file="credentials.RData")
```

### STEP 2: GETTING THE TWEETS

Once you have your authentication credentials set, you can use them to grab 
tweets from Twitter.

The next snippets of code come from my scraping_twitter.R script, which you are 
welcome to see in it's entirety on [GitHub](https://github.com/jrozra200/scraping_twitter).

``` {r logging_in, eval = FALSE}
##Authentication
load("credentials.RData") ##has my secret keys and shiz
registerTwitterOAuth(twitCred) ##logs me in

##Get the tweets about "comcast email" to work with
tweetList <- searchTwitter("comcast email", n = 1000) 

tweetList <- twListToDF(tweetList) ##converts that data we got into a data frame
```

As you can see, I used the twitteR R Package to authenticate and search Twitter. 
After getting the tweets, I converted the results to a Data Frame to make it 
easier to analyze the results.

### STEP 3: GETTING RID OF THE JUNK

Many of the tweets returned by my initial search are totally unrelated to Comcast 
Email. An example of this would be: “I am selling something random… please **email** 
me at myemailaddress@**comcast**.net”

The tweet above includes the words email and comcast, but has nothing to actually 
do with Comcast Email and the way the user feels about it, other than they use 
it for their business.

So… based on some initial, manual, analysis of the tweets, I’ve decided to pull 
those tweets with the phrases:

* “fix” AND “email” in them (in that order)
* “Comcast” AND “email” in them in that order
* “no email” in them
* Any tweet that comes from a source with “comcast” in the handle
* “Customer Service” AND “email” OR the reverse (“email” AND “Customer Service”) in them

This is done with this code:

```{r cleaning_up, eval = FALSE}
##finds the rows that have the phrase "fix ... email" in them
fixemail <- grep("(fix.*email)", tweetList$text)

##finds the rows that have the phrase "comcast ... email" in them
comcastemail <- grep("[Cc]omcast.*email", tweetList$text)

##finds the rows that have the phrase "no email" in them
noemail <- grep("no email", tweetList$text)

##finds the rows that originated from a Comcast twitter handle
comcasttweet <- grep("[Cc]omcast", tweetList$screenName)

##finds the rows related to email and customer service
custserv <- grep("[Cc]ustomer [Ss]ervice.*email|email.*[Cc]ustomer [Ss]ervice", 
                 tweetList$text)
```

After pulling out the duplicates (some tweets may fall into multiple scenarios 
from above) and ensuring they are in order (as returned initially), I assign the 
relevant tweets to a new variable with only some of the returned columns.

The returned columns are:

* text
* favorited
* favoriteCount
* replyToSN
* created
* truncated
* replyToSID
* id
* replyToUID
* statusSource
* screenName
* retweetCount
* isRetweet
* retweeted
* longitude
* latitude

All I care about are:

* text
* created
* statusSource
* screenName

This is handled through this tidbit of code:

```{r more_cleaning, eval = FALSE}
##combine all of the "good" tweets row numbers that we greped out above and 
##then sorts them and makes sure they are unique
combined <- c(fixemail, comcastemail, noemail, comcasttweet, custserv)
uvals <- unique(combined)
sorted <- sort(uvals)

##pull the row numbers that we want, and with the columns that are important to 
##us (tweet text, time of tweet, source, and username)
paredTweetList <- tweetList[sorted, c(1, 5, 10, 11)]
```

### STEP 4: CLEAN UP THE DATA AND RETURN THE RESULTS

Lastly, for this first script, I make the sources look nice, add titles, and 
return the final list (only a sample set of tweets shown):

```{r final_step, eval = FALSE}
##make the device source look nicer
paredTweetList$statusSource <- sub("<.*\">", "", paredTweetList$statusSource)
paredTweetList$statusSource <- sub("</a>", "", paredTweetList$statusSource)

##name the columns
names(paredTweetList) <- c("Tweet", "Created", "Source", "ScreenName")

paredTweetList
```

| Tweet | created | statusSource | screenName |
| ------|---------|--------------|------------|
| Dear Mark I am having problems login into my acct REDACTED@comcast.net I get no email w codes to reset my password for eddygil HELP HELP | 2014-12-23 15:44:27 | Twitter Web Client | riocauto |
| @msnbc @nbc @comcast pay @thereval who incites the murder of police officers. Time to send them a message of BOYCOTT! Tweet/email them NOW | 2014-12-23 14:52:50 | Twitter Web Client | Monty_H_Mathis |
| Comcast, I have no email. This is bad for my small business. Their response "Oh, I'm sorry for that". Problem not resolved. #comcast | 2014-12-23 09:20:14 | Twitter Web Client | mathercesul |

### CHALLENGES OBSERVED

As you can see from the output, sometimes some “junk” still gets in. Something 
I’d like to continue working on is a more reliable algorithm for identifying 
appropriate tweets. I also am worried that my choice of subjects is biasing 
the sentiment.